---
title: "Test Presentation"
author: "Kurtis Pivert"
date: "4/30/2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

***
This is a title-less slide


## Example Slide with Bullets and Lists
### Header
#### Subheading

- Bullet A
- Bullet B

1. Item number 1
2. Item number 2


## Slide with R Output

```{r cars, echo = TRUE, comment="%"}
summary(cars)
```

## Slide with Plot

```{r pressure, echo=TRUE, eval=FALSE, comment=""}
library(plotly)
xvar <- rpois(100, lambda = 50)
yvar <-rnorm(100)
zvar <- 1:100
plot_ly(x = xvar, y=yvar, z= zvar, type="scatter3d", color=yvar, mode = "markers")
```



## Next slide

Here is some text describing this model

- *Take some fake data*
- **Build a model**
- Print the model

```{r X, echo=TRUE, cache=TRUE, comment=""}
    
    x <- rnorm(100)
    y <- rnorm(100)
    model <- lm(y ~ x)
    summary(model)
    
    

``` 
