---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

This document outlines the steps to analyze anonymized data collected at 5-minute 
intervals from a personal activity monitoring device for October and 
November, 2012, including the number of steps taken.

## Loading and preprocessing the data

### Load Required Packages

First, will load required packages for this document and analysis. 

```{r load required packages, echo=TRUE}
        ## Load Required Packages
        require(dplyr)
        require(ggplot2)
        require(knitr)
        require(mice)

```


The next step is to load and read the data from the forked repository from [Github](https://github.com/kpivert/RepData_PeerAssessment1). 

```{r load and read data, echo=TRUE}
        
        ## Unzip activity.zip  
        activityData <- unzip("/Users/kpivert/RepData_PeerAssessment1/activity.zip")
        
        ## Read Data 
        stepData <- read.csv(activityData, na.strings = "NA")

```


Finally, will examine the structure of the `stepData` dataframe.

```{r structure and summary of stepData, echo=TRUE}

        str(stepData)

        summary(stepData)
        
```        

## What is mean total number of steps taken per day?

1. Make a histogram of the total number of steps taken each day. Per instructions, 
"NA" values will be omitted from analysis. 

* Will calculate total number of steps by day and produce a histogram of total 
steps by day. 

```{r total values and histogram, echo=TRUE}
        ## Calculate Total Steps By Date
        
        totalSteps <- aggregate(steps ~ date, data = stepData, sum, na.action = na.omit)
        
        ## Create Histogram Using ggplot2

        totalStepsHist <- ggplot(totalSteps, aes(steps))
        
        totalStepsHist + geom_histogram(binwidth = 500, color="grey", fill ="#003366") + 
          theme_bw()+
          ylab("Number of Days")+
          xlab("Total Steps Per Day")+
          scale_y_continuous(breaks = seq(0,8,2), 
                             labels = as.character(seq(0,8,2)))+
          ggtitle("Total Number of Steps Per Day \n Raw Dataset with NAs Removed")
        

```

2. Calculate and report the mean and median total number of steps taken per day

* Will calculate overall mean and median steps per day across 2-month data period. 

```{r mean and median for 2-month period, echo=TRUE}

      ## Mean Steps Per Day for 2-Month Period
      
      mean(stepData$steps, na.rm = TRUE)
      

      ## Median Steps Per Day for 2-Month Period
      
      median(stepData$steps, na.rm = TRUE)
      
```      


* Will also determine the mean and median total number of steps per day. Per 
instructions, will ignore missing values for this calculation. Report each set of 
values in separate tables below.

```{r mean median, echo=TRUE}

        ## Calculate Mean and Median Steps per day

        ## Mean Steps Per Day

        averageSteps <- aggregate(steps ~ date, data = stepData, mean, na.rm = TRUE)
        averageSteps$date <- strftime(averageSteps$date, format = "%B %d, %Y")
        
        ## Print Table With Mean Number of Steps Per Day
        kable(averageSteps, digits = 2, col.names = c("Date", "Mean No. of Steps"), 
              align = "c", 
              caption = "Mean Number of Steps per Day \n Raw Dataset with NAs Removed")
        
        ## Median Steps Per Day
        medianSteps <- aggregate(steps ~ date, data = stepData, median, na.rm = TRUE)
        medianSteps$date <- strftime(medianSteps$date, format = "%B %d, %Y")
        
        ## Print Table With Median Number of Steps Per Day
        kable(medianSteps, digits = 2, col.names = c("Date", "Median No. of Steps"), 
              align = "c", caption = "Median Number of Steps per Day \n Raw Dataset with NAs Removed")
        
```


## What is the average daily activity pattern?

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

* Determine average number of steps by time interval
* Create a time series plot

```{r mean daily activity patterns, echo=TRUE}

        ## Mean steps per interval
        averageInt <- aggregate(steps ~ interval, data = stepData, mean, na.rm = TRUE)

        ## Create time series plot

        averageIntPlot <- ggplot(averageInt, aes(x=interval, y=steps))
        
        averageIntPlot + geom_line(color="#003366", size = 1.5)+
          theme_bw()+
          ylab("Mean Steps per 5-Minute Interval")+
          xlab("Hours in 5-Minute Intervals")+
          ggtitle("Mean Steps Per 5-Minute Interval")+ 
          scale_x_continuous(breaks=seq(0,2000,500), labels = c("00:00", "05:00",
                "10:00", "15:00", "20:00"))


```


2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

* Determine max 5-minute interval

```{r max 5-minute interval, echo=TRUE}

        ## What is the 5-minute interval that has the largest mean number of steps
        ## across the 2-month period? 

        averageInt[which.max(averageInt$steps),1] ## Interval 835 or 8:35 a.m. 
```

## Imputing missing values

1. Calculate and report the number of `NA` values.

```{r calculate NAs, echo=TRUE}
        
        ## Determine number of NA values in dataframe
        ## apply(stepData, function(x) sum(is.na(x))) Original Code Didn't Work With             Knitr

        apply(is.na(stepData),2,sum)
```

2. Devise a strategy for filling in all of the missing values in the dataset. 
Create a new dataset that is equal to the original dataset but with the missing data filled in.
* Will use `MICE` package with default settings to complete missing values

```{r impute missing values, echo=TRUE}

        imputedStepData <- complete(mice(stepData, seed = 123))
        ## Verify no NAs remain
        sapply(imputedStepData, function(x) sum(is.na(x)))
```


3. Make a histogram of the total number of steps taken each day.

```{r histogram of total steps in imputed dataset, echo=TRUE}
        
        ## Calculate Total Steps By Date in Imputed Dataset
        
        totalImputedSteps <- aggregate(steps ~ date, data = imputedStepData, sum) 

        ## Create Histogram Using ggplot2

        totalImputedStepsHist <- ggplot(totalImputedSteps, aes(steps))
        
        totalImputedStepsHist + geom_histogram(binwidth = 500, color="grey", fill ="#003366") + 
          theme_bw()+
          ylab("Number of Days")+
          xlab("Total Steps Per Day")+
          scale_y_continuous(breaks = seq(0,8,2), 
                             labels = as.character(seq(0,8,2)))+
          ggtitle("Total Number of Steps Per Day \n Imputed Dataset")
        
```


4. Calculate and report the mean and median total number of steps taken per day. 
* First calculate mean and median steps per day

```{r mean and median of imputed dataset per day, echo=TRUE}
        
        ## Calculate Mean and Median Steps per day: Imputed Dataset

        ## Mean Steps Per Day

        averageImpSteps <- aggregate(steps ~ date, data = imputedStepData, mean)
        averageImpSteps$date <- strftime(averageImpSteps$date, format = "%B %d, %Y")
        
        ## Print Table With Mean Number of Steps Per Day: Imputed Dataset
        kable(averageImpSteps, digits = 2, col.names = c("Date", "Mean No. of Steps"),
              align = "c", 
              caption = "Mean Number of Steps per Day \n Imputed Dataset")
        
        ## Median Steps Per Day
        medianImpSteps <- aggregate(steps ~ date, data = imputedStepData, median)
        medianImpSteps$date <- strftime(medianImpSteps$date, format = "%B %d, %Y")
        
        ## Print Table With Median Number of Steps Per Day: Imputed Dataset
        kable(medianImpSteps, digits = 2, col.names = c("Date", "Median No. of Steps"), 
              align = "c", caption = "Median Number of Steps per Day \n Imputed Dataset")

```

* Then calculate overall mean and median steps per day for imputed dataset across 2-month data period. 

```{r mean and median for 2-month period imputed dataset, echo=TRUE}

      ## Mean Steps Per Day for 2-Month Period: Imputed Dataset
      
      mean(imputedStepData$steps)
      

      ## Median Steps Per Day for 2-Month Period: Imputed Dataset
      
      median(imputedStepData$steps)
      
```   

#### Do these values differ from the estimates from the first part of the assignment? 
* **Raw Datset Overall Mean:** 37.38 steps per day
* **Imputed Dataset Overall Mean:** 37.91 steps per day

*There is a minimal difference in mean steps per day between the raw and imputed dataset.*

* **Raw Datset Overall Median:** 0 steps per day
* **Imputed Dataset Overall Median:** 0 steps per day

*There is no difference in median steps per day between the raw and imputed dataset.*

#### What is the impact of imputing missing data on the estimates of the total daily number of steps?   

***Imputing missing data has no affect on estimatse of total daily number of steps.***  

## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r create weekday/weekend factor, echo=TRUE}
        
        ## Create new factor in imputed dataset
        
        ## Convert date data to date

        imputedStepData$date <- as.Date(imputedStepData$date)
        
        ## Add Days of Week Column to Dataframe
        
        imputedStepData$days <- weekdays(imputedStepData$date)
        
        ## Add New Factor
        
        mtwrf <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
        
        imputedStepData$wkDaywkEnd <- ifelse(imputedStepData$days %in% mtwrf, TRUE, FALSE) ## TRUE = Weekday, FALSE = Weekend
        
        imputedStepData$wkDaywkEnd <- factor(imputedStepData$wkDaywkEnd, labels = c("Weekend", "Weekday"))

```

2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). The plot should look something like the following, which was created using simulated data:

```{r faceted time series plot weekday vs weekend, echo=TRUE}


        ## Mean steps per interval: Imputed Dataset
        averageIntImpute <- aggregate(steps ~ interval + wkDaywkEnd, data = imputedStepData, mean)

        ## Create time series plot

        averageIntImputePlot <- ggplot(averageIntImpute, aes(x=interval, y=steps))
        
        averageIntImputePlot + geom_line(color="#003366", size = 0.9)+
          facet_grid(wkDaywkEnd ~ .)+
          theme_bw()+
          ylab("Mean Steps per 5-Minute Interval")+
          xlab("Hours in 5-Minute Intervals")+
          ggtitle("Mean Steps Per 5-Minute Interval \n Imputed Data: Weekday vs. Weekend")+ 
          scale_x_continuous(breaks=seq(0,2000,500), labels = c("00:00", "05:00",
                "10:00", "15:00", "20:00"))
        
```

